---
title: "CTA Final Essay Code"
author: "B299309"
date: "24 April 2026"
output:
  html_document:
    df_print: paged
  word_document:
    reference_docx: CTA_example.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 8,
  fig.height = 5
)
```

# Prepare Session

## Load packages
```{r}
# Load packages for data processing, text analysis, modelling, tables, and visualisation
library(data.table)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(lubridate)

library(quanteda)
library(quanteda.textmodels)
library(quanteda.textplots)
library(topicmodels)
library(tidytext)

library(knitr)
library(kableExtra)
library(forcats)
library(scales)
```

## Read data
```{r}
# Read the UK parliamentary speech dataset
data <- fread("uk_data.csv", encoding = "UTF-8")

# Convert date variable to Date format
data$date <- as.Date(data$date)

# Check the size and variable names of the dataset
dim(data)
names(data)[1:20]
```

## Create time and stage variables
```{r}
# Create half-year time periods from 2001 to 2019
data$time <- NA_character_

data$time[data$date >= as.Date("2001-01-01") & data$date <= as.Date("2001-06-30")] <- "01/1"
data$time[data$date >= as.Date("2001-07-01") & data$date <= as.Date("2001-12-31")] <- "01/2"
data$time[data$date >= as.Date("2002-01-01") & data$date <= as.Date("2002-06-30")] <- "02/1"
data$time[data$date >= as.Date("2002-07-01") & data$date <= as.Date("2002-12-31")] <- "02/2"
data$time[data$date >= as.Date("2003-01-01") & data$date <= as.Date("2003-06-30")] <- "03/1"
data$time[data$date >= as.Date("2003-07-01") & data$date <= as.Date("2003-12-31")] <- "03/2"
data$time[data$date >= as.Date("2004-01-01") & data$date <= as.Date("2004-06-30")] <- "04/1"
data$time[data$date >= as.Date("2004-07-01") & data$date <= as.Date("2004-12-31")] <- "04/2"
data$time[data$date >= as.Date("2005-01-01") & data$date <= as.Date("2005-06-30")] <- "05/1"
data$time[data$date >= as.Date("2005-07-01") & data$date <= as.Date("2005-12-31")] <- "05/2"
data$time[data$date >= as.Date("2006-01-01") & data$date <= as.Date("2006-06-30")] <- "06/1"
data$time[data$date >= as.Date("2006-07-01") & data$date <= as.Date("2006-12-31")] <- "06/2"
data$time[data$date >= as.Date("2007-01-01") & data$date <= as.Date("2007-06-30")] <- "07/1"
data$time[data$date >= as.Date("2007-07-01") & data$date <= as.Date("2007-12-31")] <- "07/2"
data$time[data$date >= as.Date("2008-01-01") & data$date <= as.Date("2008-06-30")] <- "08/1"
data$time[data$date >= as.Date("2008-07-01") & data$date <= as.Date("2008-12-31")] <- "08/2"
data$time[data$date >= as.Date("2009-01-01") & data$date <= as.Date("2009-06-30")] <- "09/1"
data$time[data$date >= as.Date("2009-07-01") & data$date <= as.Date("2009-12-31")] <- "09/2"
data$time[data$date >= as.Date("2010-01-01") & data$date <= as.Date("2010-06-30")] <- "10/1"
data$time[data$date >= as.Date("2010-07-01") & data$date <= as.Date("2010-12-31")] <- "10/2"
data$time[data$date >= as.Date("2011-01-01") & data$date <= as.Date("2011-06-30")] <- "11/1"
data$time[data$date >= as.Date("2011-07-01") & data$date <= as.Date("2011-12-31")] <- "11/2"
data$time[data$date >= as.Date("2012-01-01") & data$date <= as.Date("2012-06-30")] <- "12/1"
data$time[data$date >= as.Date("2012-07-01") & data$date <= as.Date("2012-12-31")] <- "12/2"
data$time[data$date >= as.Date("2013-01-01") & data$date <= as.Date("2013-06-30")] <- "13/1"
data$time[data$date >= as.Date("2013-07-01") & data$date <= as.Date("2013-12-31")] <- "13/2"
data$time[data$date >= as.Date("2014-01-01") & data$date <= as.Date("2014-06-30")] <- "14/1"
data$time[data$date >= as.Date("2014-07-01") & data$date <= as.Date("2014-12-31")] <- "14/2"
data$time[data$date >= as.Date("2015-01-01") & data$date <= as.Date("2015-06-30")] <- "15/1"
data$time[data$date >= as.Date("2015-07-01") & data$date <= as.Date("2015-12-31")] <- "15/2"
data$time[data$date >= as.Date("2016-01-01") & data$date <= as.Date("2016-06-30")] <- "16/1"
data$time[data$date >= as.Date("2016-07-01") & data$date <= as.Date("2016-12-31")] <- "16/2"
data$time[data$date >= as.Date("2017-01-01") & data$date <= as.Date("2017-06-30")] <- "17/1"
data$time[data$date >= as.Date("2017-07-01") & data$date <= as.Date("2017-12-31")] <- "17/2"
data$time[data$date >= as.Date("2018-01-01") & data$date <= as.Date("2018-06-30")] <- "18/1"
data$time[data$date >= as.Date("2018-07-01") & data$date <= as.Date("2018-12-31")] <- "18/2"
data$time[data$date >= as.Date("2019-01-01") & data$date <= as.Date("2019-06-30")] <- "19/1"
data$time[data$date >= as.Date("2019-07-01") & data$date <= as.Date("2019-12-31")] <- "19/2"

# Create an alternative time variable without slash
data$time2 <- str_replace(data$time, "/", "_")

# Create parliamentary stage variable
data$stage <- 0
data$stage[data$m_questions == 1] <- 1
data$stage[data$u_questions == 1] <- 2
data$stage[data$queen_debate_others == 1] <- 3
data$stage[data$queen_debate_day1 == 1] <- 4
data$stage[data$pm_questions == 1] <- 5

# Convert stage into labelled factor
data$stage <- factor(
  data$stage,
  levels = c(0, 1, 2, 3, 4, 5),
  labels = c("Other", "Ministerial Questions", "Urgent Questions",
             "Queen Debate Other", "Queen Debate Day 1", "PM Questions")
)

# Check missing time values and stage distribution
sum(is.na(data$time))
table(data$stage, useNA = "ifany")
```

## Subset Labour PMQs
```{r}
# Select Labour speeches in Prime Minister's Questions
pmq_labour <- data %>%
  filter(pm_questions == 1, party == "Labour")

# Check sample size and government status distribution
nrow(pmq_labour)
table(pmq_labour$government, useNA = "ifany")

# Keep variables needed for the analysis
pmq_labour <- pmq_labour %>%
  select(id_speech, date, time, time2, party, government, text)

head(pmq_labour)
```

## Check government and opposition periods
```{r}
# Check whether Labour speeches are divided into government and opposition periods
pmq_labour %>%
  mutate(year = as.numeric(format(date, "%Y"))) %>%
  group_by(year) %>%
  summarise(
    gov_share = mean(government),
    n = n(),
    .groups = "drop"
  ) %>%
  kable(
    digits = 3,
    caption = "Share of Labour PMQ speeches in government by year"
  ) %>%
  kable_styling(full_width = FALSE)
```

## Plot government status over time
```{r plot-government-status, fig.width=8, fig.height=4.8, dpi=300, fig.cap="Figure A1. Share of Labour speeches in government by year"}
# Visualise the government status of Labour speeches over time
pmq_labour %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarise(gov_share = mean(government), .groups = "drop") %>%
  ggplot(aes(x = year, y = gov_share)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Share of Labour speeches in government by year",
    y = "Government proportion",
    x = "Year"
  ) +
  theme_minimal()
```

The figure shows that from 2001 to 2019 Labour was in government continuously until around 2010, after which it moved into opposition and remained there for the rest of the period.

# Topic Modelling (LDA)

## Aggregate texts by half-year period
```{r}
# Aggregate Labour PMQ speeches into half-year time periods
pmq_labour_time <- pmq_labour %>%
  group_by(time) %>%
  summarise(
    text = paste(text, collapse = " "),
    mean_date = mean(date),
    .groups = "drop"
  ) %>%
  mutate(
    period = ifelse(mean_date < as.Date("2010-05-01"),
                    "Government", "Opposition")
  )
```

## Create corpus for LDA
```{r}
# Create a corpus from time-aggregated Labour PMQ texts
corp_time <- corpus(pmq_labour_time, text_field = "text")

# Tokenise texts and remove punctuation, symbols, numbers, and URLs
toks_time <- tokens(
  corp_time,
  remove_punct = TRUE,
  remove_symbols = TRUE,
  remove_numbers = TRUE,
  remove_url = TRUE
) %>%
  tokens_tolower()

# Define custom stopwords for parliamentary and generic terms
custom_stopwords <- c(
  stopwords("en"),

  # parliamentary procedural words
  "hon", "member", "members", "minister", "ministers", "secretary",
  "gentleman", "gentlelady", "lady", "house", "prime", "government",
  "mr", "mrs", "speaker", "right", "friend", "friends",
  "colleagues", "official", "engagements", "meeting", "meetings",

  # generic discourse fillers
  "will", "can", "could", "would", "may", "also", "just",
  "now", "make", "made", "say", "says", "said", "going",
  "go", "get", "got", "let", "today", "year", "years",
  "time", "people", "one", "know", "think", "take", "many",
  "want", "way", "week", "last", "first", "new", "well",

  # common PMQ routine words
  "morning", "afternoon", "evening", "wednesday", "thursday", "monday",
  "duty", "duties", "shall", "later", "question", "answer",

  # additional vague high-frequency words
  "important", "sure", "agree", "however", "point", "clear",
  "next", "done", "action", "report", "case", "matter",
  "whether", "forward", "believe", "hope", "understand",
  "aware", "yesterday", "others", "part", "whole"
)

# Remove stopwords
toks_time <- tokens_remove(toks_time, pattern = custom_stopwords)

# Keep tokens with at least three characters
toks_time <- tokens_keep(toks_time, min_nchar = 3)

# Add unigrams and bigrams
toks_time <- tokens_ngrams(toks_time, n = 1:2)

# Create document-feature matrix
dfm_time <- dfm(toks_time)

# Inspect most frequent features before further cleaning
topfeatures(dfm_time, 50)
```

## Additional corpus cleaning for LDA
```{r}
# Remove additional high-frequency words with limited semantic meaning
extra_stopwords_clean <- c(
  "need", "help", "give", "put", "look", "must", "see",
  "every", "back", "two", "tell", "course", "given"
)

toks_time <- tokens_remove(toks_time, pattern = extra_stopwords_clean)

# Recreate dfm after additional cleaning
dfm_time <- dfm(toks_time)

# Inspect frequent terms again
topfeatures(dfm_time, 50)
```

## Initial LDA model comparison
```{r}
# Trim the dfm before topic modelling
dfm_time_trim <- dfm_trim(
  dfm_time,
  min_docfreq = 0.05,
  max_docfreq = 0.85,
  docfreq_type = "prop"
)

# Convert quanteda dfm into topicmodels format
dtm_time <- convert(dfm_time_trim, to = "topicmodels")

# Estimate 3-, 4-, and 5-topic models for comparison
set.seed(1234)

lda_time_3 <- LDA(dtm_time, k = 3, method = "VEM")
lda_time_4 <- LDA(dtm_time, k = 4, method = "VEM")
lda_time_5 <- LDA(dtm_time, k = 5, method = "VEM")

# Inspect top terms from each model
terms(lda_time_3, 15)
terms(lda_time_4, 15)
terms(lda_time_5, 15)
```

## Final LDA cleaning and model estimation
```{r}
# Remove a second set of additional terms after inspecting initial model results
extra_stopwords_2 <- c(
  "cent", "per_cent", "deputy",
  "addition", "additional", "amount",
  "entirely", "precisely", "although",
  "promised", "promises", "broken",
  "opposed", "grateful", "wrote",
  "touch", "rule", "good", "number",
  "since", "past", "result", "future",
  "behaviour"
)

toks_time <- tokens_remove(toks_time, pattern = extra_stopwords_2)

# Recreate and trim dfm for the final LDA models
dfm_time <- dfm(toks_time)

dfm_time_trim <- dfm_trim(
  dfm_time,
  min_docfreq = 0.05,
  max_docfreq = 0.85,
  docfreq_type = "prop"
)

# Convert to topicmodels format
dtm_time <- convert(dfm_time_trim, to = "topicmodels")

# Estimate final 3-, 4-, and 5-topic models
set.seed(1234)

lda_time_3 <- LDA(dtm_time, k = 3, method = "VEM", control = list(seed = 1234))
lda_time_4 <- LDA(dtm_time, k = 4, method = "VEM", control = list(seed = 1234))
lda_time_5 <- LDA(dtm_time, k = 5, method = "VEM", control = list(seed = 1234))

# Inspect topic terms
terms(lda_time_3, 15)
terms(lda_time_4, 15)
terms(lda_time_5, 15)
```

The k=3 model produced three clearly interpretable and distinct topics, corresponding to (1) welfare and domestic policy, (2) economic conditions and global issues, and (3) war and security. The keywords within each topic showed strong internal consistency and meaningful substantive interpretation.

In contrast, the k=4 and k=5 models resulted in more fragmented topic structures. Several topics overlapped in content (e.g., multiple topics containing war-related terms such as “iraq,” “afghanistan,” and “troops”), and some topics lacked clear thematic coherence. This made interpretation less straightforward.

Based on this comparison, the k=3 model was selected for subsequent analysis.

## Extract LDA(k=3) beta values
```{r}
# Extract beta values from the final 3-topic LDA model
beta_time_3 <- tidy(lda_time_3, matrix = "beta")

# Keep the top 15 terms for each topic
beta_top15 <- beta_time_3 %>%
  group_by(topic) %>%
  slice_max(beta, n = 15) %>%
  ungroup()
```

## Table of topic keywords
```{r}
# Display top 15 terms for each topic
beta_top15 %>%
  arrange(topic, desc(beta)) %>%
  kable(
    digits = 4,
    caption = "Top 15 terms for each topic in the LDA model"
  ) %>%
  kable_styling(full_width = FALSE)
```

## Plot LDA topic keywords
```{r plot-lda-topic-keywords, fig.width=12, fig.height=6.8, dpi=500, fig.cap="Figure 1. Top 15 terms for each LDA topic (k=3)"}
# Visualise the highest-beta terms in each topic

topic_labels <- c(
  "1" = "1. Welfare and domestic policy",
  "2" = "2. Economic conditions and global issues",
  "3" = "3. War and security"
)

beta_top15 %>%
  mutate(
    topic = factor(topic),
    topic_label = recode(as.character(topic), !!!topic_labels),
    term = reorder_within(term, beta, topic_label)
  ) %>%
  ggplot(aes(x = beta, y = term, fill = topic)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic_label, scales = "free", ncol = 3) +
  scale_y_reordered() +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 12) +
  labs(
    title = "Top 15 terms by topic (LDA, k = 3)",
    x = "Beta (term probability)",
    y = NULL
  ) +
  theme(
    strip.text = element_text(size = 8, face = "bold"),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(size = 8),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    panel.spacing = unit(1.2, "lines")
  )
```

The figure shows three main topics in Labour’s discourse: welfare and domestic policy (e.g. brexit, unemployment, social care), economic and global issues (e.g. afghanistan, banks, recession), and war and security (e.g. iraq, terrorism, troops).

## Extract topic distribution values
```{r}
# Extract topic probabilities from the final 3-topic LDA model
gamma_time <- posterior(lda_time_3)$topics

# Convert gamma matrix into a data frame
topic_gamma_df <- as.data.frame(gamma_time)
colnames(topic_gamma_df) <- paste0("Topic_", 1:3)

# Combine gamma values with the time-aggregated Labour PMQ data
topic_gamma_df <- cbind(pmq_labour_time, topic_gamma_df)
```

## Convert gamma values into long format
```{r}
# Convert topic probabilities into long format
topic_gamma_long <- topic_gamma_df %>%
  pivot_longer(
    cols = starts_with("Topic_"),
    names_to = "topic",
    values_to = "gamma"
  ) %>%
  mutate(
    topic_label = recode(
      topic,
      "Topic_1" = "Welfare and domestic policy",
      "Topic_2" = "Economic conditions and global issues",
      "Topic_3" = "War and security"
    )
  )
```

## Table of topic prevalence over time
```{r}
# Create a wide table of gamma values by time period
topic_gamma_wide <- topic_gamma_long %>%
  ungroup() %>%
  select(time, topic_label, gamma) %>%
  pivot_wider(names_from = topic_label, values_from = gamma)

# Keep periods in chronological order
topic_gamma_wide <- topic_gamma_wide %>%
  mutate(time = factor(time, levels = unique(pmq_labour_time$time))) %>%
  arrange(time)

# Display gamma values
topic_gamma_wide %>%
  kable(
    digits = 3,
    caption = "Topic prevalence over time (gamma values)"
  ) %>%
  kable_styling(full_width = FALSE)
```

## Plot topic prevalence over time
```{r plot-topic-prevalence, fig.width=12, fig.height=6.8, dpi=500, fig.cap="Figure 2. Topic prevalence over time in Labour PMQs (LDA, k = 3)"}
# Plot topic probabilities across half-year periods
ggplot(topic_gamma_long, aes(x = time, y = gamma, group = topic_label, color = topic_label)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_vline(xintercept = which(unique(pmq_labour_time$time) == "10/1"),
             linetype = "dashed") +
  theme_minimal(base_size = 12) +
  labs(
    title = "Topic prevalence over time in Labour PMQs (LDA, k = 3)",
    x = "Time",
    y = "Topic probability",
    color = "Topic"
  ) +
  theme(
    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 11, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 11),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    legend.position = "bottom"
  )
```

The figure shows a clear shift in dominant topics over time: discourse is initially dominated by war and security, then shifts to economic issues, and after 2010 becomes dominated by welfare and domestic policy.

## LDA robustness check with different random seeds
```{r}
# Re-estimate the 3-topic LDA model using different random seeds
set.seed(1234)
lda_3_a <- LDA(dtm_time, k = 3, method = "VEM")

set.seed(5678)
lda_3_b <- LDA(dtm_time, k = 3, method = "VEM")

set.seed(9999)
lda_3_c <- LDA(dtm_time, k = 3, method = "VEM")

# Compare top terms across different model runs
terms(lda_3_a, 15)
terms(lda_3_b, 15)
terms(lda_3_c, 15)
```

## External topic validation using top_topic
```{r}
# Create a fuller Labour PMQ dataset including the supervised top_topic variable
pmq_labour_full <- data %>%
  filter(pm_questions == 1, party == "Labour") %>%
  select(id_speech, date, time, time2, party, government, text, top_topic)

# Display the distribution of top_topic categories
pmq_labour_full %>%
  count(top_topic, sort = TRUE) %>%
  kable(caption = "Distribution of top_topic in Labour PMQs") %>%
  kable_styling(full_width = FALSE)
```

# Discourse Position (Wordfish)

## Create Wordfish corpus
```{r}
# Create a corpus for Wordfish using speech-level Labour PMQ texts
corp_wf <- corpus(pmq_labour, text_field = "text")

# Add time as a document-level variable
docvars(corp_wf, "time") <- pmq_labour$time

# Tokenise texts for Wordfish
toks_wf <- tokens(
  corp_wf,
  remove_punct = TRUE,
  remove_symbols = TRUE,
  remove_numbers = TRUE,
  remove_url = TRUE
) %>%
  tokens_tolower()

# Define stopwords for Wordfish
wf_stopwords <- c(
  stopwords("en"),
  "hon", "member", "members", "minister", "ministers", "secretary",
  "gentleman", "gentlelady", "lady", "house", "prime", "government",
  "mr", "mrs", "speaker", "right", "friend", "friends"
)

# Remove stopwords
toks_wf <- tokens_remove(toks_wf, pattern = wf_stopwords)

# Create dfm
dfm_wf_raw <- dfm(toks_wf)

# Group dfm by half-year period
dfm_wf_time <- dfm_group(dfm_wf_raw, groups = time)

# Trim low-frequency features
dfm_wf_time <- dfm_trim(
  dfm_wf_time,
  min_docfreq = 2,
  docfreq_type = "count"
)

# Check dimensions
dim(dfm_wf_time)
```

## Estimate Wordfish model
```{r}
# Estimate Wordfish model
set.seed(123L)

wf_time <- textmodel_wordfish(dfm_wf_time)
```

## Plot document positions on the Wordfish scale
```{r plot-wordfish-scale, fig.width=12, fig.height=6.8, dpi=500, fig.cap="Figure A2. Document positions on the Wordfish scale"}
textplot_scale1d(wf_time)
```

The figure shows that Labour’s discourse positions are spread across the scale. Some periods are clustered on the right, while others appear on the left, suggesting substantial variation in how discourse is structured.

## Extract Wordfish feature scores
```{r}
# Extract Wordfish feature beta values

features <- wf_time[["features"]]
betas <- wf_time[["beta"]]

feat_betas <- data.frame(
  feature = features,
  beta = as.numeric(betas)
)

head(feat_betas, 10)
```


## Positive Wordfish features
```{r}
# Display features at the positive end of the Wordfish dimension
feat_betas %>%
  arrange(desc(beta)) %>%
  slice_head(n = 20) %>%
  kable(digits = 3, caption = "Top positive Wordfish features") %>%
  kable_styling(full_width = FALSE)
```

## Negative Wordfish features
```{r}
# Display features at the negative end of the Wordfish dimension
feat_betas %>%
  arrange(beta) %>%
  slice_head(n = 20) %>%
  kable(digits = 3, caption = "Top negative Wordfish features") %>%
  kable_styling(full_width = FALSE)
```

## Plot positive and negative features
```{r plot-wordfish-features, fig.width=12, fig.height=6.8, dpi=500, fig.cap="Figure 3. Most distinctive Wordfish features at both ends of the scale"}

# Prepare positive and negative Wordfish features
top_pos <- feat_betas %>%
  arrange(desc(beta)) %>%
  slice_head(n = 20) %>%
  mutate(direction = "Positive")

top_neg <- feat_betas %>%
  arrange(beta) %>%
  slice_head(n = 20) %>%
  mutate(direction = "Negative")

plot_data <- bind_rows(top_pos, top_neg)

# Flip negative side for mirror effect
plot_data <- plot_data %>%
  mutate(beta_plot = ifelse(direction == "Negative", -abs(beta), abs(beta)))

# Plot distinctive Wordfish features
ggplot(plot_data, aes(x = beta_plot, y = reorder(feature, beta_plot), fill = direction)) +
  geom_col(show.legend = TRUE) +
  theme_minimal() +
  labs(
    title = "Most distinctive Wordfish features",
    x = "Beta (direction of discourse)",
    y = NULL,
    fill = NULL
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.y = element_text(size = 10)
  )
```

The figure shows a clear contrast between two sides of the Wordfish dimension: one side is characterised by international security terms (e.g. iraq, saddam, disarm), while the other is dominated by domestic political issues (e.g. brexit, austerity, grenfell).

## Extract Wordfish positions over time
```{r}
# Extract estimated Wordfish positions
positions <- data.frame(
  time = rownames(dfm_wf_time),
  theta = wf_time$theta
)

# Add government/opposition period labels
pmq_labour_time$period <- ifelse(
  pmq_labour_time$mean_date < as.Date("2010-05-01"),
  "Government",
  "Opposition"
)

positions <- merge(positions, pmq_labour_time[, c("time", "period")])

# Keep time in correct chronological order
positions$time <- factor(positions$time, levels = unique(pmq_labour_time$time))

positions <- positions %>%
  arrange(time)
```

## Plot Wordfish estimated positions over time
```{r plot-wordfish-position, fig.width=12, fig.height=6.8, dpi=500, fig.cap="Figure 4. Discourse position over time based on the Wordfish model"}
ggplot(positions, aes(x = time, y = theta, color = period, group = 1)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_vline(xintercept = which(levels(positions$time) == "10/1"),
             linetype = "dashed") +
  theme_minimal(base_size = 12) +
  labs(
    title = "Discourse position over time (Wordfish)",
    x = "Time",
    y = "Estimated position",
    color = "Period"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 11, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 11),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    legend.position = "bottom"
  )
```

The figure shows a clear shift in discourse position over time, with a sharp change around 2010 from a higher (positive) position during government to a lower (negative) position during opposition.

# Sentiment Analysis (AFINN and Bing)

## Prepare tokens for sentiment analysis
```{r}
# Convert time-aggregated texts into tidy token format
tidy_tokens <- pmq_labour_time %>%
  unnest_tokens(word, text)
```

## Sentiment analysis using AFINN lexicon
```{r}
# Load AFINN sentiment lexicon
afinn <- get_sentiments("afinn")

# Match tokens with AFINN lexicon
sentiment_afinn_df <- tidy_tokens %>%
  inner_join(afinn, by = "word")

# Calculate average AFINN sentiment score by time period
sent_afinn_time <- sentiment_afinn_df %>%
  group_by(time) %>%
  summarise(
    score = mean(value),
    .groups = "drop"
  ) %>%
  mutate(lexicon = "AFINN")
```

## Plot AFINN sentiment over time
```{r plot-sentiment-afinn, fig.width=12, fig.height=6.8, dpi=500, fig.cap="Figure 5. Sentiment over time based on the AFINN lexicon"}
ggplot(sent_afinn_time, aes(x = time, y = score, group = 1)) +
  geom_line(linewidth = 1, color = "#F8766D") +
  geom_point(size = 2, color = "#F8766D") +
  geom_vline(xintercept = which(sent_afinn_time$time == "10/1"),
             linetype = "dashed") +
  theme_minimal(base_size = 12) +
  labs(
    title = "Sentiment over time using AFINN lexicon",
    x = "Time",
    y = "Average sentiment score"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 11, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 11)
  )
```

The figure shows a clear decline in sentiment over time, with a sharp drop around 2010 from positive to more negative values.

## Sentiment analysis using Bing lexicon
```{r}
# Load Bing sentiment lexicon
bing <- get_sentiments("bing")

# Match tokens with Bing lexicon
sentiment_bing_df <- tidy_tokens %>%
  inner_join(bing, by = "word")

# Calculate sentiment score by time period
sent_bing_time <- sentiment_bing_df %>%
  count(time, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(
    score = (positive - negative) / (positive + negative + 1),
    lexicon = "Bing"
  )
```

## Plot Bing sentiment over time
```{r plot-sentiment-bing, fig.width=12, fig.height=6.8, dpi=500, fig.cap="Figure A3. Sentiment over time based on the Bing lexicon"}
ggplot(sent_bing_time, aes(x = time, y = score, group = 1)) +
  geom_line(linewidth = 1, color = "#00BFC4") +
  geom_point(size = 2, color = "#00BFC4") +
  geom_vline(xintercept = which(levels(factor(sent_bing_time$time)) == "10/1"),
             linetype = "dashed") +
  theme_minimal(base_size = 12) +
  labs(
    title = "Sentiment over time using Bing lexicon",
    x = "Time",
    y = "Sentiment score"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 11, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 11)
  )
```

The figure shows a general decline in sentiment over time, with a noticeable drop around 2010 and mostly lower or negative values afterward.

## Sentiment robustness check
```{r}
# Combine AFINN and Bing results for robustness comparison
sent_robustness <- bind_rows(
  sent_afinn_time,
  sent_bing_time %>% select(time, score, lexicon)
)

# Keep time periods in chronological order
sent_robustness <- sent_robustness %>%
  mutate(time = factor(time, levels = unique(pmq_labour_time$time)))
```

## Plot sentiment robustness check
```{r plot-sentiment-robustness, fig.width=12, fig.height=6.8, dpi=500, fig.cap="Figure 6. Sentiment robustness check comparing AFINN and Bing lexicons"}
ggplot(sent_robustness, aes(x = time, y = score, color = lexicon, group = lexicon)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.8) +
  geom_vline(xintercept = which(levels(sent_robustness$time) == "10/1"),
             linetype = "dashed") +
  theme_minimal(base_size = 12) +
  labs(
    title = "Sentiment robustness check: AFINN and Bing",
    x = "Time",
    y = "Sentiment score",
    color = "Lexicon"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 11, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 11),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    legend.position = "right"
  )
```

The figure shows that both AFINN and Bing follow a very similar trend over time, with sentiment declining around 2010 and remaining lower afterward.








